---
title: "Individual Assignment - Data Analyzes"
output: html_notebook
---

What is the number of fatalities in Israel in the year 2001? In the following we want to try to answer this question by analyzing the UCDP conflict dataset file. Each line (each entry) of the dataset is one conflict. (For more information what a conflict is: Sundberg, Ralph & Erik Melander, 2013)

About the dataset:
"Thus, the dataset is constructed in such a way as maximize the comparability and consistency across time and space, and provide a globally consistent image of the
phenomenon of organized violence.In effect, the goal of UCDP GED is not to present the most complete and accurate image of a certain conflict at a certain point in time, but rather be a tool for the global understanding of subnational conflict patterns and trends." - (Croicu, Mihai & Sundberg, 2016)

The data has a lot of variables, amongst them the country, the year of the conflict, the best estimated guess for total fatalities (aggregated for both sides), type of conflict, region, etc.

For our goal it shall be enough to look at the country, year, best estimated aggregated guess, and possibly some more variables of future interest, which are filtered out during the R processing.

Methods:
First, load the tidyverse for additional funtions and read the csv file we created out of the json file by using Python.


```{r}
#loading tidyverse
  library(tidyverse)
  conflicts <- read_csv("../data/israel_data.csv")

```

Next, we filter out all relevant columns for the research question. First we need all entries for Israel from the whole dataset.
```{r}
#First, we select from our data only the entries for the country Israel.
israel_data <- conflicts %>%
  filter (Country == 'Israel')
```
Then, we need all entries for the year 2001.
```{r}
#Second, we select from all Israel entries the ones for the year 2001
israel_2001 <- israel_data %>%
    filter (Year == 2001)
```
Now that we have the data we need, we can give a result of the fatalities in Israel in the year 2001. Let us create a bargraph for that...because we can!
```{r}
# Bar-plot of year and amount of fatalities (best guessed)
ggplot (data = israel_2001) +
  geom_col (mapping = aes(x = Year, y = best_fatality_guess),  fill = "blue")  +
  xlab("Year") +
  ylab("Fatalities in 100") +
  ggtitle("Amount of fatalities in Israel in 2001") +

  ggsave('israel_fatal_2001.pdf')

```
Is it not beautiful?
Anyway, we have a graph now! But we would still like to know the actual number.

```{r}
israel_2001 %>%
  summarize(total_fatalities = sum(best_fatality_guess))
```
The total sum of fatalities in Israel in 2001 is 355.


Now, we want to compare the total fatalities in Israel over years. First, we start comparing it to 2002, the following year of the 9/11 attacks. The attacks having happened quite late in 2001, we would expect an increase in conflicts for the following year at least.

```{r}
# Filtering out the year 2002 or 2001
israel_2002_01 <- israel_data %>% 
  filter (Year == 2002 | Year == 2001)
```

```{r}
# grouping it by year
israel_02_01_grouped <- israel_2002_01 %>%
  group_by(Year) %>%
  summarize(sum=sum(best_fatality_guess, na.rm=T))
  
```

```{r}
# Bar-plot for 2001 and 2002 and amount of fatalities (best guessed)
ggplot (data = israel_02_01_grouped) +
  geom_col (mapping = aes(x = Year, y = sum),
            fill = "blue"
            )  +
  xlab("Year") +
  scale_x_continuous(breaks = c(2001, 2002)) +
  ylab("Fatalities in 100") +
  ggtitle("Amount of fatalities in Israel in 2001 & 2002") +
  
  ggsave('israel_fatal_2001_02.pdf')
```
```{r}
israel_2002_01 %>%
  summarize(total_fatalities = sum(best_fatality_guess))
```
The total sum of fatalities in Israel in 2001 is 355. Therefore, the total fatalities for the year 2002 equals to 1235-355 = 880.
(cheap, maybe, but works at the moment)


```{r}
#comparison for Israel over several years, full data set.
israel_years <- israel_data %>%
  filter (Year >= 1989, Year <= 2015)
```

```{r}
# Bar-plot for amount of fatalities for Israel over years for the full data set 1989 - 2015 (26 years)
ggplot (data = israel_years) +
  geom_col (mapping = aes(x = Year, y = best_fatality_guess),  fill = "blue")  +
  xlab("Year") +
  ylab("Fatalities in 100") +
  ggtitle("Amount of fatalities in Israel between 1990 and 2010") +
  
  ggsave('israel_over_years.pdf')
```
```{r}
israel_years %>%
  summarize(fatalities = sum(best_fatality_guess))
```
```{r}
# Selecting columns relevant for future questions
israel_years_rel <- israel_years %>%
  group_by(Year) %>% 
  select(Country, Year, best_fatality_guess, type_of_violence)
```


Bibliography:
Sundberg, Ralph, and Erik Melander, 2013, “Introducing the UCDP Georeferenced Event Dataset”,
Journal of Peace Research, vol.50, no.4, 523-532

Croicu, Mihai and Ralph Sundberg, 2016, “UCDP GED Codebook version 5.0”, Department of
Peace and Conflict Research, Uppsala University
